---
title: "flow_rate"
author: "Erin Ristig"
date: "5/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load the required packages
library(DeclareDesign)
library(knitr)
library(ggplot2)
library(grid)
library(gridExtra)
library(dplyr)
library(tidyverse)
library(kableExtra)
```

##Measure:
Flow rate (meters/second) measurements collected at established sites

##Target Population: 
Every meter of total waterway length in the Sandy River Basin - including all tributataries to the Sandy River. Estimated to be 165,000 meters. 

## Hypothetical Population

```{r pop}
#Description of Hypothetical Population: 

set.seed(6)
population <- declare_population(
  stream_meters = add_level(N = 165000, # total number of meters
                            accessible = c(rep(0,85000), rep(1, 80000)), # 0 is location isn't easily accessible from road (roughly a third); 1 is easily accessible (other two thirds)
  flow_rate = if_else(accessible == 1, rnorm(n = 165000, mean = 200, sd = 120), rnorm(n = 165000, mean= 600, sd = 150))
)) # associated flow; if accesible by road will likely be in an area with less steep terrain - this means the flow rate will likely be lower in these areas as there will be less elevation change. 

# 0 = location not easily accessible from a road, 1 = location easily accessible from a road

#mean and sd in meters/second, flow_rate is flow rate at given site along river

pop <- population()



my_estimand <- declare_estimands(mean(flow_rate),
                                 label = "Ybar")

plot <- ggplot(pop, aes(x=flow_rate)) + 
  geom_histogram(color="black", fill="white")

plot


```
##Challenges of drawing a representative sample: 
Given limited staff capacity, flow rate measurements are limited in terms of number of sites and frequency. Stream volume, natural obstacles (ie eddies), human caused obstacles (ie log jams), elevation change, etc. all vary greatly throughout the Sandy River Basin. A representative sample of this measurement would include flow rate measurements from sites encompassing the full range of possibilities and combinations of these characteristics within the basin. This would require establishing sites at all headwater tributaries in the upper basin, and establishing sites at equal length intervals down all waterways until the mouth of the Sandy River. Given that not all sites along the river are accessible and time and budget are limited it is not feasible to measure flow rate along the entirety of the river. Site selection will be biased towards locations easily accesible from a road. Areas of the river with steeper terrain are likely to be less accesible as there will be fewer roads. Steep terrain areas are difficult to construct roads on and tend to be generally less developed. The flow rate of the river in areas of steep terrain is likely to be faster as the elevation change is more dramatic causing waterfalls, cascades, etc. This means that steeper areas of the basin, with faster flow rates will be harder to access and therefore under-represented.

This is addressed in the code in the stratification and the weighting of the stratification. Explanation from Jenny in slack
Not including reporting because every sight we're going to will have water

##Reporting probabilities

```{r diff-reporting, echo=FALSE}
# Reporting probabilities

#reporting <- declare_assignment(blocks=accessible,
                              #  assignment_variable = "R", #site sampled
                               # block_prob=c(0.1,0.6)) 

#pop <- reporting(pop)

```

## Sampling Procedure
With two types of sites - those easily accessible from a road, and those not easily accessible - it makes the most sense to use a disproportionate strata sampling procedure. It isn't feasible to sample the less accessible sites more intensively (hence a bias in the first place), but using the strata_weighted_mean() function we can weight the strata means proportionately to the strata size of the population. 

```{r diagnosis, cache=TRUE}
# Sampling procedure

sampling <- declare_sampling(strata = accessible,
                             strata_n = c(9,11))

strata_weighted_mean <- function(data){
  data.frame( 
  estimator_label = "strata_w_mean",
  estimand_label = "Ybar",
  n = nrow(data),
  stringsAsFactors = FALSE,
 
  estimate = data %>% 
   group_by(accessible) %>% 
  summarise(mean=mean(flow_rate)) %>%
  mutate(prop=c(0.5152,0.4848)) %>%
  mutate(sub.mean=mean*prop) %>% pull(sub.mean) %>% 
  sum())
} 



answer <- declare_estimator(
 handler = tidy_estimator(strata_weighted_mean),
 estimand = my_estimand)

design <- population + my_estimand + reporting + sampling + answer

diagnosis <- diagnose_design(design, sims = 200)

diagnosis$diagnosands_df[,c(4,5,12,14)] %>%
 kable()

```



