---
title: "De-list endangered salmon and steelhead of the Lower Columbia River Basin"
subtitle: "Effectiveness of guided site visits"
author: "Minnie"
output: beamer_presentation
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load, echo=FALSE}
# Load the required packages
library(DeclareDesign)
library(knitr)
library(ggplot2)
library(grid)
library(gridExtra)
library(dplyr)
library(kableExtra)
```

## Measurement Goal & Measure

**Goal**: Test effectiveness of guided site visits in increasing program awareness and making good impressions on visitors.

" Please rate your experience today [with X guide] on the following (Likert) scale:"

a. Very Satisfied
b. Somewhat Satisfied
c. Neutral
d. Somewhat Dissatisfied
e. Very Dissatisfied

*Note: present in horizontal line to prevent order bias

## Source & Rationale & Unit

*Source*: Each visitor will be asked to complete the survey following the guided site visit.

*Rationale*: The site visits are provided in order to provide potential donors and collaborators with first-hand experience of the program elements and impacts. The assumption we are testing is whether site visits truly increase visitor knowledge and leave guests with a positive impression. This measure of visit satisfaction will later be correlated with actual future engagement.

*Unit*: Individual survey respondent


## Responsibility & Frequency

*Responsibility*: The survey should be administered by a staff member (possibly administrator) other than the tour guide, or via email. The survey should be partially anonymous - guests should not feel uncomfortable providing honest feedback, but it will be helpful to include some identifying information that will allow other measures (which organization they represent and the $ amount of the donation or hours of work on a collaboration) to be correlated with tour satisfaction.

*Frequency*: Every individual should be asked to take the survey, since each could represent a different donor organization. Additionally, if there are multiple tour guides, it would be important to determine which/whose strategy is most effective. 


## Declaring the population

Administrative data shows that 20 tours are led throughout the year, with 10 individuals are on each tour, for a total of 200 visitors.

Organizations       Visitors               Surveys Completed
-----------        ------------------      ------------------
Will Donate              160                      100%
Will Not Donate          40                       50%


## Target Population & Challenges

*Target Population*: All visitors who go on a site tour

*Challenge of drawing a representative sample*: Some individuals may refuse to complete the survey, and this could be correlated with their level of (dis)satisfaction with the visit - this would bias the survey towards positive feedback (sampling bias).

*Sampling procedure*: Request every visitor complete the survey at the end of their site visit.

## Declare Population

```{r population}
set.seed(228)

#Minnie: Create a population of 200 visitors, of which 80% will end up donating. Then use whether or not visitors are donors to determine their level of satisfaction. This means that visitors' donor status is directly correlated to whether or not they were satisfied with the tour (the assumption we are testing).

population <- declare_population(
  visitors = add_level(N=200, donor = sample(c(rep(0,40),rep(1,160))),
                       satisfied=correlate(given = donor, rho = 0.8, draw_binary, prob = 0.5)))

#Minnie: To make satisfaction a binary condition, a survey response of "neutral" or more negative is considered "dissatisfied" and both "somewhat satisfied" and "very satisfied" are considered "satisfied".
#Minnie:rho = 0.8 --> donor status highly correlated with being satisfied

pop <- population()
pop.vector <- c(40,160)

kable(table(pop$donor,pop$satisfied)) %>% 
  add_header_above(c("donor"=1,"satisfied"=2))

my_estimand <- declare_estimands(mean(pop$satisfied),
                                 label = "Ybar")
#Minnie: Estimand is the mean level of satisfaction among the population
#my_estimand(pop)

```

## Declare Response Bias

```{r reporting and sampling, echo=FALSE}

reporting <- declare_assignment(blocks=donor,
                  assignment_variable = "A", 
                  block_prob=c(0.5,1.0))

#Minnie: Donors will all complete the survey, while half of non-donors will refuse to complete it --> donors are more likely to "Answer"

rep <- reporting(pop)
table(rep$donor,rep$A)


```


## Declare Sampling

```{r sampling and diagnosis, cache=TRUE}

sampling <- declare_sampling(strata=donor,
                             strata_n = c(9,1)) #Minnie: this appears to be the lowest possible sampling number/distribution but the bias is reduced by weighting towards non-donor surveys
samp <- sampling(rep)

strata_weighted_mean <- function(data){
  data.frame(  
  estimator_label = "strata_w_mean",
  estimand_label = "Ybar",
  n = nrow(data),
  stringsAsFactors = FALSE,
  
  estimate = data %>% filter(A==1) %>%
    group_by(donor) %>%
    summarise(estimate=mean(satisfied)) %>%
    mutate(prop=c(0.2,0.8)) %>%
    mutate(sub.mean=estimate*prop) %>% pull(sub.mean) %>%
    sum())
}

answer <- declare_estimator(
  handler = tidy_estimator(strata_weighted_mean), estimand = my_estimand)

design <- population + reporting + sampling + my_estimand + answer
diagnosis <- diagnose_design(design, sims = 1000)

diagnosis$diagnosands_df[,c(4,5,12,14)] %>%
  kable()

```



